Strong coupling between scales in a multi-scalar model of urban dynamics

### 2022-10-26 Docs/Papers/Multiscale/paper/paper.tex
figs_paper

\author*[1,2,3,4]{\pfx{Dr} \fnm{Juste} \sur{Raimbault} \dgr{PhD}}\email{juste.raimbault@ign.fr}
% orcid 0000-0003-0768-9480

\affil*[1]{\orgdiv{LASTIG}, \orgname{Univ. Gustave Eiffel, IGN-ENSG}} %, \orgaddress{\street{73 avenue de Paris}, \city{Saint-Mand{\'e}}, \postcode{93160}, \country{France}}
\affil[2]{\orgdiv{Center for Advanced Spatial Analysis}, \orgname{UCL}} % , \orgaddress{\street{90 Tottenham Court Road}, \city{London}, \postcode{}, \country{UK}}
\affil[3]{\orgdiv{UPS CNRS 3611 ISC-PIF}, \orgname{CNRS}} % , \orgaddress{\street{Place Nationale}, \city{Paris}, \postcode{75013}, \country{France}}
\affil[4]{\orgdiv{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s}, \orgname{Universit{\'e} Paris 1}} % , \orgaddress{\street{Place Nationale}, \city{Paris}, \postcode{75013}, \country{France}}

@@

%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
 \includegraphics[width=\linewidth]{figures/multiscale_morph.pdf}
 \caption{Schematic representation of the model architecture. Urban areas with their own urban form are inserted into the macroscopic spatial interaction model, and strong coupling is achieved through top-down and bottom-up feedbacks.\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%%%%%

@@

Mesoscopic grids population grids are evolved separately within each urban area, using the updated morphogenesis model parameters. We follow here the same exact steps described in \cite{raimbault2018calibration}, with parameters $N_G^{(i)}$ total population increase, $\beta_i$ diffusion level, $\alpha_i$ aggregation level, $t_m$ time steps and $n_d$ number of diffusions. The model adds iteratively population to grid cells, by distributing it with a probability following a preferential attachment to population, and diffusing it.

@@

The Table~\ref{tab:parameters} summarises model parameters. For clarity, we do not include parameters linked to synthetic model setup, and distinguish them by the scale to which they are attached.
% TODO ref table beginning section

@@

Model behavior is characterized using the following indicators: at the  macroscopic scale: distributions of population, accessibilities, and centralities (summarized by average, hierarchy, entropy), following \cite{raimbault2020unveiling}. At the mesoscopic scale, urban form is captured following the indicators used by \cite{raimbault2018calibration}, i.e. Moran index, average distance between individuals, hierarchy of population distribution, and entropy of population distribution.

@@

Model implementation is done under some performance constraints: $N$ mesoscopic morphogenesis models must be simulated in parallel at each macroscopic time step. On the contrary, macroscopic interactions are efficient to implement as they are based on matrices computations. The model is implemented in \texttt{scala} and integrated into a broader library for spatial sensitivity analysis and dynamical models for settlement systems, namely the \texttt{spatialdata} library \cite{raimbault2020scala}. This library in particular already includes implementations of the models used in \cite{raimbault2020indirect} and \cite{raimbault2018calibration}, but also of other urban interaction models such as \cite{favaro2011gibrat}. The relative large number of parameters and output indicators imposes the use of appropriate model exploration methods to get a grasp of model behavior. Therefore, the parameter space is explored with the OpenMOLE model exploration software \cite{reuillon2013openmole}. Standalone source code for the model is available on an open git repository as \cite{urbangrowthmodel}. An other open repository for the analysis of results is also available as \cite{urbangrowth}. Simulation data used in this paper is available on the dataverse repository as \cite{dataverse}.

@@

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/allindics_hist.png}
	\caption{Statistical distribution of indicators for a sample of 10 parameter values.\label{fig:fig2}}
\end{figure}

@@

We then turn to a grid exploration of the parameter space. %precise successive experiments, step values and varied parameters
Results show an important role of the strong coupling between scales. This is shown for example by the variation of macroscopic indicators trajectories when switching from a ``transit-oriented development'' scenario (negative feedback of population growth on diffusion $\delta \beta$) to a ``sprawl'' scenario (positive feedback $\delta \beta$) as shown in Fig.~\ref{fig:fig4} below.

\begin{figure}[H]
	\includegraphics[width=\linewidth]{figures/deltaHierarchymacroPopulations-mesoTimeSteps_colorMacroInteractionDecay_facetmesoMacroDecayUpdateMax-macroMesoBetaUpdateMax.png}
	\caption{Macroscopic hierarchy of populations as a function of mesoscopic time steps $t_m$, for different values of interaction decay $d_G$ (colour), and of feedback parameters $\delta d$ (columns) and $\delta \beta$ (rows).\label{fig:fig4}}
\end{figure}

We show in Fig.~\ref{fig:fig3} that both macroscopic and mesoscopic indicators have a U-shaped behavior as a function of the bottom-up feedback parameter $\delta d$. In terms of macroscopic trajectories, we see that hierarchies of accessibilities, closeness centrality and population are maximal in negative value when the mechanism is deactivated, what means that the feedback process moderates the hierarchisation of the global urban system. For the urban form, this feedback also mitigates the aggregation of population.

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/onefactor_allindics_mesoMacroDecayUpdateMax_errorbars.png}
	\caption{Variations of indicators as a function of $\delta d$, when all other parameters are fixed to their nominal value.\label{fig:fig3}}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/deltamesoMorans-macroMesoAlphaUpdateMax_colormacroMesoBetaUpdateMax_facetmesoMacroCongestionCost-mesoMacroDecayUpdateMax_mesoBeta0_11.png}
	\caption{Relative evolution of average mescoscopic Moran, as a function of feedback parameter $\delta \beta$, for different values of feedback parameter $\delta \alpha$ (colour), and of feedback parameters $\delta d$ (columns) and $\lambda$ (rows).\label{fig:fig5}}
\end{figure}

@@

When investigating population hierarchy the mesoscopic scale, i.e. inequalities of population distribution within the urban area, we find in Fig.~\ref{fig:fig6} a U-shaped behavior as a function of $\delta \beta$ for negative $\delta \alpha$ values, but has a plateau for positive values of $\delta \alpha$, showing also an interaction effect between policies. In terms of hierarchy, a maximal negative value may be wanted to limit inequalities (slopes are negative), and this optimum is reached for a slightly positive value of $\delta \beta$ (left columns in Fig.~\ref{fig:fig6}). This means that sprawl may be needed in some configurations to optimise this objective.

@@

This model is only a first structural sketch with very restrictive assumption, in particular regarding the downward and upward feedbacks on submodel parameters. There may be no link between urban form and global insertion, or it may be due to other processes, be expressed as an other functional form. An important stage before shifting to robust knowledge will consist in (i) reviewing and making a typology of such potential processes across scales; (ii) including most in a multi-modeling fashion to compare possible concurrent mechanisms.

>>>>

\author{}
%\author*[1,2,3,4]{\pfx{Dr} \fnm{Juste} \sur{Raimbault} }\email{juste.raimbault@ign.fr}% \dgr{PhD} % orcid 0000-0003-0768-9480

%\affil*[1]{\orgdiv{LASTIG}, \orgname{Univ. Gustave Eiffel, IGN-ENSG}} %, \orgaddress{\street{73 avenue de Paris}, \city{Saint-Mand{\'e}}, \postcode{93160}, \country{France}}
%\affil[2]{\orgdiv{Center for Advanced Spatial Analysis}, \orgname{UCL}} % , \orgaddress{\street{90 Tottenham Court Road}, \city{London}, \postcode{}, \country{UK}}
%\affil[3]{\orgdiv{UPS CNRS 3611 ISC-PIF}, \orgname{CNRS}} % , \orgaddress{\street{Place Nationale}, \city{Paris}, \postcode{75013}, \country{France}}
%\affil[4]{\orgdiv{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s}, \orgname{Universit{\'e} Paris 1}} % , \orgaddress{\street{Place Nationale}, \city{Paris}, \postcode{75013}, \country{France}}

@@

%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
 \includegraphics[width=\linewidth]{figures/Fig1.jpg}
 \caption{Schematic representation of the model architecture. Urban areas with their own urban form are inserted into the macroscopic spatial interaction model, and strong coupling is achieved through top-down and bottom-up feedbacks.\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%%%%%

@@

Mesoscopic grids population grids are evolved separately within each urban area, using the updated morphogenesis model parameters. We follow here the same exact steps described in \cite{raimbault2018calibration}, with parameters $N_G^{(i)}$ total population increase, $\beta_i$ diffusion level, $\alpha_i$ aggregation level, $t_m$ time steps and $n_d$ number of diffusions. The model iteratively adds population to grid cells by distributing it with a probability following a preferential attachment to population, and then diffuses it in space.

@@

Model parameters are summarised in Table~\ref{tab:parameters}. For clarity, we do not include parameters linked to synthetic model setup, and distinguish them by the scale to which they correspond.

@@

Model behaviour is characterised using the following indicators: at the  macroscopic scale: distributions of population, accessibilities, and centralities (summarized by average, hierarchy, entropy), following \cite{raimbault2020unveiling}. At the mesoscopic scale, urban form is captured following the indicators used by \cite{raimbault2018calibration}, i.e. Moran index, average distance between individuals, hierarchy of population distribution, and entropy of population distribution.

@@

Model implementation has to meet some performance constraints: $N$ mesoscopic morphogenesis models must be simulated in parallel at each macroscopic time step. On the contrary, macroscopic interactions are efficient to implement as they are based on matrices computations. The model is implemented in \texttt{scala} and integrated into a broader library for spatial sensitivity analysis and dynamical models for settlement systems, namely the \texttt{spatialdata} library \cite{raimbault2020scala}. This library in particular already includes implementations of the models used in \cite{raimbault2020indirect} and \cite{raimbault2018calibration}, but also of other urban interaction models such as \cite{favaro2011gibrat}. The relative large number of parameters and output indicators imposes the use of appropriate model exploration methods to get a grasp of model behavior. Therefore, the parameter space is explored with the OpenMOLE model exploration software \cite{reuillon2013openmole}. Standalone source code for the model is available on an open git repository as \cite{urbangrowthmodel}. An other open repository for the analysis of results is also available as \cite{urbangrowth}. Simulation data used in this paper is available on the dataverse repository as \cite{dataverse}.

@@

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig2.jpg}
	\caption{Statistical distribution of indicators for a sample of 10 parameter values.\label{fig:fig2}}
\end{figure}

@@

We then turn to a grid exploration of the parameter space. %precise successive experiments, step values and varied parameters
Results show an important role of the strong coupling between scales. This is shown for example by the variation of macroscopic indicators trajectories when switching from a ``transit-oriented development'' scenario (negative feedback of population growth on diffusion $\delta \beta$) to a ``sprawl'' scenario (positive feedback $\delta \beta$) as shown in Fig.~\ref{fig:fig3} below.

\begin{figure}[H]
	\includegraphics[width=\linewidth]{figures/Fig3.jpg}
	\caption{Macroscopic hierarchy of populations as a function of mesoscopic time steps $t_m$, for different values of interaction decay $d_G$ (colour), and of feedback parameters $\delta d$ (columns) and $\delta \beta$ (rows).\label{fig:fig3}}
\end{figure}

We show in Fig.~\ref{fig:fig4} that both macroscopic and mesoscopic indicators have a U-shaped behaviour as a function of the bottom-up feedback parameter $\delta d$. In terms of macroscopic trajectories, we see that hierarchies of accessibilities, closeness centrality and population are maximal in negative value when the mechanism is deactivated, what means that the feedback process moderates the hierarchisation of the global urban system. For the urban form, this feedback also mitigates the aggregation of population.

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig4.jpg}
	\caption{Variations of indicators as a function of $\delta d$, when all other parameters are fixed to their nominal value.\label{fig:fig4}}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig5.jpg}
	\caption{Relative evolution of average mescoscopic Moran, as a function of feedback parameter $\delta \beta$, for different values of feedback parameter $\delta \alpha$ (colour), and of feedback parameters $\delta d$ (columns) and $\lambda$ (rows).\label{fig:fig5}}
\end{figure}

@@

When investigating population hierarchy the mesoscopic scale, i.e. inequalities of population distribution within the urban area, we find in Fig.~\ref{fig:fig6} a U-shaped behaviour as a function of $\delta \beta$ for negative $\delta \alpha$ values, but has a plateau for positive values of $\delta \alpha$, showing also an interaction effect between policies. In terms of hierarchy, a maximal negative value may be wanted to limit inequalities (slopes are negative), and this optimum is reached for a slightly positive value of $\delta \beta$ (left columns in Fig.~\ref{fig:fig6}). This means that sprawl may be needed in some configurations to optimise this objective.

@@

This model is only a first structural sketch with very restrictive assumption, in particular regarding the downward and upward feedbacks on submodel parameters. There may be no link between urban form and global insertion, or it may be due to other processes, be expressed as an other functional form. An important stage before shifting to robust knowledge will consist in (i) reviewing and making a typology of such potential processes across scales; (ii) including most in a multi-modelling fashion to compare possible concurrent mechanisms.
